---
title: "HIV_2009_ANALYSIS_REPORT_NIGERIA"
author: "*By Qozeem Uthman*"
format: 
  dashboard
theme: lumen
---


```{r setup, include=FALSE, message=FALSE}
#Load packages
if(!require(pacman)) install.packages("pacman")
pacman::p_load(here, tidyverse,
               ggplot2, plotly, reactable,
               bslib, shiny, gapminder,
               bsicons, reactable, htmltools,
               highcharter, flexdashboard,
               rnaturalearth, plotly, countrycode,
               ggtext, sf, flextable)

```

```{r}
import_data<-read_csv(here("data/hiv_incidence_pop.csv"))
import_data1 <- import_data %>%
  filter(year %in% 2005:2009)

data2<-read_csv(here("data/hiv_prev_pop.csv"))
import_data2 <- data2 %>%
  filter(year %in% 2005:2009)

data3<-read_csv(here("data/newly_hiv_infected_number_all_ages.csv"))
import_data3 <- data3 %>%
  filter(country == "Nigeria") %>%
  select(`2005`:`2011`) %>%
  pivot_longer(cols = `2005`:`2011`, names_to = "Year", values_to = "Figure") %>%
  arrange(desc(Year)) %>% reactable()

data4<-read_csv(here("data/people_living_with_hiv_number_all_ages.csv"))
import_data4 <- data4 %>%
  filter(country == "Nigeria") %>%
  select(`2005`:`2011`) %>%
  pivot_longer(cols = `2005`:`2011`, names_to = "Year", values_to = "Figure") %>%
  arrange(desc(Year)) %>% reactable()


```


```{r}
# Filter the data for Nigeria
nigeria_data <- import_data1 %>% filter(country == "Nigeria")

# Calculate HIV incidence rate per year for Nigeria
nigeria_data <- nigeria_data %>%
  mutate(incidence_rate = (new_cases / population) * 100000)

# Display the incidence rate per year
nigeria_incidence <- nigeria_data %>% select(year, incidence_rate)

Incidence_Rate<-nigeria_incidence%>%filter(year==2009)%>% select(incidence_rate) %>% pull(incidence_rate)%>%
  round(1)

# Visualization: Interactive plot of new cases over the years for Nigeria
# Define custom colors

# Example custom colors
mycolors <- c("#add8e6", "#4A5A73", "#FFB6C1", "#dddddd",  "#E6E6FA")

new_cases_plot <- ggplot(nigeria_data, aes(x = year, y = new_cases)) +
  geom_line() +
  geom_point(size = 1) +
  labs(title = "",
       x = "Year",
       y = "New Cases",
       color = "Year") +
  theme_minimal()

# Convert ggplot object to plotly for interactivity
plot1<-ggplotly(new_cases_plot)



# Visualization: Interactive plot of population over the years for Nigeria
population_plot <- ggplot(nigeria_data, aes(x = year, y = population)) +
  geom_line() +
  geom_point(size = 1) +
  labs(title = "",
       x = "Year",
       y = "Population",
       color = "Year") +
  theme_minimal()

# Convert ggplot object to plotly for interactivity
plot2<-ggplotly(population_plot)

```


```{r}
# Filter the data for Nigeria
nigeria_data2 <- import_data2 %>% filter(country == "Nigeria")

# Calculate HIV prevalence rate per year for Nigeria
nigeria_data2 <- nigeria_data2 %>%
  mutate(prevalence_rate = (total_cases / population) * 100000)

# Display the prevalence rate per year
nigeria_prevalence <- nigeria_data2 %>% select(year, prevalence_rate)

Prevalence_Rate <-HIV_prevelance<-nigeria_prevalence%>%
  filter(year==2009)%>% 
  select(prevalence_rate) %>% pull(prevalence_rate)%>%
  round(1)

# Visualization: Interactive plot of total cases over the years for Nigeria
total_cases_plot <- ggplot(nigeria_data2, aes(x = year, y = total_cases)) +
  geom_line() +
  geom_point() +
  labs(title = "",
       x = "Year",
       y = "Total Cases") +
  theme_minimal()

plot3<-ggplotly(total_cases_plot)


# Calculate total cases for each year
total_cases_year <- nigeria_data2 %>%
  group_by(year) %>%
  summarise(total_cases = sum(total_cases))

pie_chart <- plot_ly(labels = total_cases_year$year,
                     values = total_cases_year$total_cases,
                     type = "pie",
                     marker = list(colors = mycolors),
                     textinfo = "label+percent",
                     hoverinfo = "text+percent",
                     textposition = "inside",
                     hole = 0.6) %>%
  layout(showlegend = FALSE)

plot4<-pie_chart

```

```{r}
nig_data1 <- import_data %>% filter(country == "Nigeria")
nig_data2 <- data2 %>% filter(country == "Nigeria")

# Combine the datasets
combined_data <- nig_data1 %>%
    select(country, year, new_cases, population) %>%
    inner_join(nig_data2 %>% select(country, year, total_cases, population), 
               by = c("country", "year", "population"))

# Select only the columns you need
final_data <- combined_data %>%
    select(country, year, population, new_cases, total_cases)
```


# DASHBOARD

## Row 1 {height=20%}
```{r}
value_box(
  title = "Incidence Rate",
  value = Incidence_Rate,
  showcase = bsicons::bs_icon("save-fill"),
  theme = value_box_theme(bg = "#4A5A73"))
```

```{r}
value_box(
  title = "Prevalence Rate",
  value = Prevalence_Rate,
  showcase = bsicons::bs_icon("infinity"),
  theme = value_box_theme(bg = "#343a40"))
```

```{r}
value_box(
  title = "Newly HIV Infected",
  value = "360K",
  showcase = bsicons::bs_icon("search"),
  theme = value_box_theme(bg = "#add8e6")
)
```

```{r}
value_box(
  title = "People Living with HIV",
  value = "3.2M",
  showcase = bsicons::bs_icon("people-fill"),
  theme = value_box_theme(bg = "#E6E6FA")
)
```

## Row 2 {height=80%}

### {width=35%}

```{r title = "HIV Incidence in Nigeria"}
plot1
```

### {width=30%}

```{r title = "Distribution of HIV Cases by Year in Nigeria"}
plot4
```
### {width=35%}


```{r title = "HIV Prevalence in Nigeria"}
plot3
```


# NIGERIA HIV SCENE

## Row 1 {height=30%}

```{r title = "Population in Nigeria"}
plot2
```

## Row 2 {height=25%}

### {width=20%}

```{r title = "Newly_HIV_infected"}
import_data3
```

### {width=25%}

```{r title = "People_living_with_HIV"}
import_data4
```

## {.sidebar width=20%}

Future research topics related to Human Immunodeficiency Virus (HIV), we can explore several important areas based on the available data provided:

### Introduction to HIV

Human Immunodeficiency Virus (HIV) is a virus that attacks the body’s immune system, specifically the CD4 cells (T cells), which are crucial for fighting infections. If untreated, HIV can lead to Acquired Immunodeficiency Syndrome (AIDS). Unlike some other viruses, the human body cannot get rid of HIV completely, so once someone has HIV, they have it for life.

### Symptoms and Causes

HIV is primarily spread through unprotected sex, sharing needles, and from mother to child during childbirth or breastfeeding. Initial symptoms can include fever, chills, rash, night sweats, muscle aches, sore throat, fatigue, swollen lymph nodes, and mouth ulcers. These symptoms can appear within 2-4 weeks after exposure and may last for a few weeks. As the virus progresses, it can lead to more severe symptoms and opportunistic infections.

### Important Things to Know

1. **Diagnosis and Treatment**: Early diagnosis through testing is crucial. Antiretroviral therapy (ART) is the standard treatment and, if taken consistently, can control the virus, allowing individuals to live long, healthy lives and reducing the risk of transmission.
2. **Prevention**: Preventive measures include using condoms, pre-exposure prophylaxis (PrEP), and not sharing needles. Education and awareness are key in preventing the spread of HIV.

### Future Research and Recommendations

Future research focuses on finding a cure, developing effective vaccines, and creating novel antiretroviral therapies. Promising areas include exploring gene editing technologies and latency reversal agents. Addressing the intersection of HIV with non-communicable diseases and aging, as well as overcoming stigma and improving healthcare access, are also critical  .

---

**References**:

1. UNAIDS. (2023). [Global HIV & AIDS statistics](https://www.unaids.org/en/resources/fact-sheet).
2. CDC. (2022). [HIV Basics](https://www.cdc.gov/hiv/basics/index.html).


# DOWNLOAD DATA

The subset of the raw data used in this dashboard is shown below and can be downloaded as a csv

```{r}
htmltools::browsable(
  tagList(
    reactable(final_data,
              elementId = "final_data",
              searchable = TRUE,
              filterable = TRUE),
    tags$button("Download as CSV",
                onclick = "Reactable.downloadDataCSV('final_data')")
  )
)

```